### 2022-12-30 submission.tex
update gp question

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage[margin=2cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage[export]{adjustbox}
\usepackage{indentfirst}
\usepackage{wrapfig}
\usepackage{listings}
\usepackage{color}
\usepackage{enumerate}
\usepackage{amssymb, bm}
\usepackage{amsmath}
\usepackage{csvsimple}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{lscape}
\newcommand\tab[1][1cm]{\hspace*{#1}}
\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}
\lstset{ %
    backgroundcolor=\color{gray!10!white},
  basicstyle=\tiny, %footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  deletekeywords={...},            % if you want to delete keywords from the given language
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  language=Python,                 % the language of the code
  morekeywords={*,...},           % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=10pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{mygray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=1,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mymauve},     % string literal style
  tabsize=1,	                   % sets default tabsize to 2 spaces
  title=\lstname                   % show the filename of files included with \lstinputlisting; also try caption instead of title
}
\graphicspath{ }
\usetikzlibrary{arrows}

\title{\textbf{COMP0085 Summative Assignment}}
%\author{Jian Shu (James) Wu \\ }
\date{Jan 4, 2023}

\begin{document}
\maketitle
\section*{Question 1}

\subsection*{(a)} The directed acyclic graph:

\begin{center}
\begin{tikzpicture}[node distance={15mm}, main/.style = {draw, circle}]
    \node[main] (A) {$A$};
    \node[main] (B) [below left of=A] {$B$};
    \node[main] (C) [below right of=A] {$C$};
    \node[main] (E) [below right of=C] {$E$};
    \node[main] (D) [below left of=C] {$D$};
    \node[main] (H) [below of=E] {$H$};
    \node[main] (G) [right of=E] {$G$};
    \node[main] (I) [below left of=H] {$I$};
    \node[main] (F) [above left of=I] {$F$};

    \draw[-latex] (A) -- (B);
    \draw[-latex] (A) -- (C);
    \draw[-latex] (B) -- (F);
    \draw[-latex] (C) -- (D);
    \draw[-latex] (C) -- (E);
    \draw[-latex] (G) -- (H);
    \draw[-latex] (D) -- (F);
    \draw[-latex] (E) -- (H);
    \draw[-latex] (F) -- (I);
    \draw[-latex] (H) -- (I);
\end{tikzpicture}
\end{center}

\subsection*{(b)} The moralised graph:

\begin{center}
\begin{tikzpicture}[node distance={15mm}, main/.style = {draw, circle}]
    \node[main] (A) {$A$};
    \node[main] (B) [below left of=A] {$B$};
    \node[main] (C) [below right of=A] {$C$};
    \node[main] (E) [below right of=C] {$E$};
    \node[main] (D) [below left of=C] {$D$};
    \node[main] (H) [below of=E] {$H$};
    \node[main] (G) [right of=E] {$G$};
    \node[main] (I) [below left of=H] {$I$};
    \node[main] (F) [above left of=I] {$F$};

    \draw (A) -- (B);
    \draw (A) -- (C);
    \draw (B) -- (F);
    \draw (B) -- (D);
    \draw (C) -- (D);
    \draw (C) -- (E);
    \draw (E) -- (G);
    \draw (G) -- (H);
    \draw (D) -- (F);
    \draw (E) -- (H);
    \draw (F) -- (H);
    \draw (F) -- (I);
    \draw (H) -- (I);
\end{tikzpicture}
\end{center}

\newpage

An effective triangulation:

%\begin{center}
%\begin{tikzpicture}[node distance={15mm}, main/.style = {draw, circle}]
%    \node[main] (A) {$A$};
%    \node[main] (B) [below left of=A] {$B$};
%    \node[main] (C) [below right of=A] {$C$};
%    \node[main] (E) [below right of=C] {$E$};
%    \node[main] (D) [below left of=C] {$D$};
%    \node[main] (H) [below of=E] {$H$};
%    \node[main] (G) [right of=E] {$G$};
%    \node[main] (I) [below left of=H] {$I$};
%    \node[main] (F) [above left of=I] {$F$};
%
%    \draw (A) -- (B);
%    \draw (A) -- (C);
%    \draw (B) -- (F);
%    \draw (B) -- (D);
%    \draw (C) -- (D);
%    \draw (C) -- (E);
%    \draw (E) -- (G);
%    \draw (G) -- (H);
%    \draw (D) -- (F);
%    \draw (E) -- (H);
%    \draw (F) -- (H);
%    \draw (F) -- (I);
%    \draw (H) -- (I);
%
%    \draw[dashed] (I) -- (E);
%    \draw[dashed] (F) -- (E);
%    \draw[dashed] (C) -- (F);
%    \draw[dashed] (B) -- (D);
%    \draw[dashed] (B) -- (C);
%
%\end{tikzpicture}
%\end{center}

\begin{center}
\begin{tikzpicture}[node distance={15mm}, main/.style = {draw, circle}]
    \node[main] (A) {$A$};
    \node[main] (B) [below left of=A] {$B$};
    \node[main] (C) [below right of=A] {$C$};
    \node[main] (E) [below right of=C] {$E$};
    \node[main] (D) [below left of=C] {$D$};
    \node[main] (H) [below of=E] {$H$};
    \node[main] (G) [right of=E] {$G$};
    \node[main] (I) [below left of=H] {$I$};
    \node[main] (F) [above left of=I] {$F$};

    \draw (A) -- (B);
    \draw (A) -- (C);
    \draw (B) -- (F);
    \draw (B) -- (D);
    \draw (C) -- (D);
    \draw (C) -- (E);
    \draw (E) -- (G);
    \draw (G) -- (H);
    \draw (D) -- (F);
    \draw (E) -- (H);
    \draw (F) -- (H);
    \draw (F) -- (I);
    \draw (H) -- (I);

    \draw[dashed] (C) -- (H);
    \draw[dashed] (C) -- (F);
    \draw[dashed] (B) -- (D);
    \draw[dashed] (B) -- (C);

\end{tikzpicture}
\end{center}

where the dashed lines are edges added to triangulate the moralised graph.

\newpage

The resulting junction tree:

%\begin{center}
%\begin{tikzpicture}[node distance={15mm}, main/.style = {draw, circle}, square/.style={regular polygon,regular polygon sides=4}]
%    \node[main] (ABC) {$ABC$};
%    \node[square] (BC) [square,draw] [below of=ABC] {$BC$};
%    \node[main] (BCDF) [below of=BC] {$BCDF$};
%    \node[square] (CF) [square,draw] [below of=BCDF] {$CF$};
%    \node[main] (CEF) [below of=CF] {$CEF$};
%    \node[square] (EF) [square,draw] [below of=CEF] {$EF$};
%    \node[main] (EFHI) [below of=EF] {$EFHI$};
%    \node[square] (EH) [square,draw] [below of=EFHI] {$EH$};
%    \node[main] (EGH) [below of=EH] {$EGH$};
%    \draw (ABC) -- (BC);
%    \draw (BC) -- (BCDF);
%    \draw (BCDF) -- (CF);
%    \draw (CF) -- (CEF);
%    \draw (CEF) -- (EF);
%    \draw (EF) -- (EFHI);
%    \draw (EFHI) -- (EH);
%    \draw (EH) -- (EGH);
%\end{tikzpicture}
%\end{center}

\begin{center}
\begin{tikzpicture}[node distance={25mm}, main/.style = {draw, circle}, square/.style={regular polygon,regular polygon sides=4}]
    \node[main] (ABC) {$ABC$};
    \node[main] (BCDF) [below of=ABC] {$BCDF$};
    \node[main] (CFH) [below of=BCDF] {$CFH$};
    \node[main] (FHI) [below left of=CFH] {$FHI$};
    \node[main] (CEH) [below right of=CFH] {$CEH$};
    \node[main] (EGH) [below of=CEH] {$EGH$};
    \draw (ABC) -- (BCDF);
    \draw (BCDF) -- (CFH);
    \draw (CFH) -- (FHI);
    \draw (CFH)  -- (CEH);
    \draw (EGH)  -- (CEH);
\end{tikzpicture}
\end{center}

where the circular nodes are cliques.

\newpage

The junction tree redrawn as a factor graph:

\begin{center}
\begin{tikzpicture}[node distance={15mm}, main/.style = {draw, circle}, square/.style={regular polygon,regular polygon sides=4}]
    \node[main] (ABC) {$ABC$};
    \node[square] (BC) [square,draw] [below of=ABC] {$BC$};
    \node[main] (BCDF) [below of=BC] {$BCDF$};
    \node[square] (CF) [square,draw] [below of=BCDF] {$CF$};
    \node[main] (CFH) [below of=CF] {$CFH$};
    \node[square] (FH) [square,draw] [below left of=CFH] {$FH$};
    \node[main] (FHI) [below left of=FH] {$FHI$};
    \node[square] (CH) [square,draw] [below right of=CFH] {$CH$};
    \node[main] (CEH) [below right of=CH] {$CEH$};
    \node[square] (EH) [square,draw] [below of=CEH] {$EH$};
    \node[main] (EGH) [below of=EH] {$EGH$};
    \draw (ABC) -- (BC);
    \draw (BC) -- (BCDF);
    \draw (BCDF) -- (CF);
    \draw (CF) -- (CFH);
    \draw (CFH) -- (FH);
    \draw (FH) -- (FHI);
    \draw (CFH) -- (CH);
    \draw (CH) -- (CEH);
    \draw (CEH) -- (EH);
    \draw (EGH) -- (EH);
\end{tikzpicture}
\end{center}

where the circular nodes are cliques and the square nodes are separators/factors.

\newpage

\subsection*{(c)}

\begin{center}
\begin{tikzpicture}[node distance={15mm}, main/.style = {draw, circle}]
    \node[main] (A) [fill=black!20] {$A$};
    \node[main] (B) [below left of=A] {$B$};
    \node[main] (C) [below right of=A] {$C$};
    \node[main] (E) [below right of=C, fill=black!20] {$E$};
    \node[main] (D) [below left of=C, fill=black!20] {$D$};
    \node[main] (H) [below of=E, fill=black!20] {$H$};
    \node[main] (G) [right of=E] {$G$};
    \node[main] (I) [below left of=H] {$I$};
    \node[main] (F) [above left of=I, fill=black!20] {$F$};

    \draw[-latex] (A) -- (B);
    \draw[-latex] (A) -- (C);
    \draw[-latex] (B) -- (F);
    \draw[-latex] (C) -- (D);
    \draw[-latex] (C) -- (E);
    \draw[-latex] (G) -- (H);
    \draw[-latex] (D) -- (F);
    \draw[-latex] (E) -- (H);
    \draw[-latex] (F) -- (I);
    \draw[-latex] (H) -- (I);
\end{tikzpicture}
\end{center}

The set $\{A, D, E, F, H\}$ is a non-unique smallest set of molecules such that if the concentrations of the species within the set are known, the concentrations of the others $\{B, C, G, I\}$ would all be independent (conditioned on the measured ones).

\subsection*{(d)}

\newpage

\subsection*{(e)}

\newpage
\section*{Question 2}

\subsection*{(a)}

We want the posterior mean and covariance over $a$ and $b$.
Defining a weight vector $\textbf{w}$:
\[\textbf{w} = \begin{bmatrix}
                 a \\
                 b
         \end{bmatrix}\]

Our distribution for $\textbf{w}$:
\[P(\textbf{w}) = \mathcal{N} \left(
\begin{bmatrix}
                 \mu_a \\
                 \mu_b
         \end{bmatrix} ,
\begin{bmatrix}
                 \sigma_a^2 & 0 \\
                 0 & \sigma_b^2
         \end{bmatrix}
\right)
  = \mathcal{N}(\mu_{\textbf{w}}, \Sigma_{\textbf{w}})
\]

Moreover, for our data $\mathcal{D} = \{\textbf{X}, \textbf{Y}\}$:
\[P(\mathcal{D} | \textbf{w}) = \mathcal{N} \left( \textbf{Y} - \textbf{w}^T \textbf{X}, \sigma^2 \textbf{I}
\right)
\]

where $\textbf{X} =  \begin{bmatrix}
                 t_1 & t_2 \cdots t_N \\
                 1 & 1 \cdots 1
         \end{bmatrix} \in  \mathbb{R}^{2 \times N}$ and $\textbf{Y} \in \mathbb{R}^{1 \times N}$.

Knowing:
\[P(\textbf{w} | \mathcal{D}) \propto P(\mathcal{D} | \textbf{w}) P(\textbf{w})\]

we can substitute the above distributions:
\[P(\textbf{w} | \mathcal{D}) \propto
  \exp \left(\frac{-1}{2 \sigma^2} \left( \textbf{Y} - \textbf{w}^T \textbf{X}\right) \left( \textbf{Y} - \textbf{w}^T \textbf{X}\right)^T \right)
\exp \left(\frac{-1}{2} \left( \textbf{w} -\mu_{\textbf{w}}\right)^T \Sigma_{\textbf{w}}^{-1} \left( \textbf{w} -\mu_{\textbf{w}}\right)\right)
\]

expanding:
\[\log P(\textbf{w} | \mathcal{D}) \propto
  \frac{-1}{2} \left( \frac{\textbf{Y} \textbf{Y}^T}{\sigma^2} - 2\textbf{w}^T \frac{\textbf{X}\textbf{Y}^T}{\sigma^2} + \textbf{w}^T \frac{\textbf{X} \textbf{X}^T}{\sigma^2} \textbf{w} +  \textbf{w}^T \Sigma_{\textbf{w}}^{-1}  \textbf{w} - 2\textbf{w}^T \Sigma_{\textbf{w}}^{-1}\mu_{\textbf{w}} + \mu_{\textbf{w}}^T\Sigma_{\textbf{w}}^{-1}\mu_{\textbf{w}} \right)
\]

collecting $\textbf{w}$ terms:
\[\log P(\textbf{w} | \mathcal{D}) \propto
  \frac{-1}{2} \left( \textbf{w}^T \left(\frac{\textbf{X} \textbf{X}^T}{\sigma^2} + \Sigma_{\textbf{w}}^{-1} \right)  \textbf{w} - 2\textbf{w}^T \left(\frac{\textbf{X}\textbf{Y}^T}{\sigma^2} + \Sigma_{\textbf{w}}^{-1} \mu_{\textbf{w}}\right)  \right)
\]

Knowing that the posterior $P(\textbf{w} | \mathcal{D})$ will be Gaussian with mean $\bar{\mu}_w$ and covariance $\bar{\Sigma}_w$, we can see that expanding the exponent component would have the form:
\[\left( \textbf{w} - \bar{\mu}_w \right)^T \bar{\Sigma}_w^{-1} \left( \textbf{w} - \bar{\mu}_w \right) = \textbf{w}^T \bar{\Sigma}_w^{-1} \textbf{w} -2 \textbf{w}^T \bar{\Sigma}_w^{-1} \bar{\mu}_w + \bar{\mu}_w^T \bar{\Sigma}_w^{-1} \bar{\mu}_w\]

Thus we can identify the posterior covariance:
\[\bar{\Sigma}_w = \left(\frac{\textbf{X} \textbf{X}^T}{\sigma^2} + \Sigma_{\textbf{w}}^{-1} \right)^{-1}\]

and the posterior mean:
\[\bar{\mu}_w = \bar{\Sigma}_w \left(\frac{\textbf{X}\textbf{Y}^T}{\sigma^2} + \Sigma_{\textbf{w}}^{-1} \mu_{\textbf{w}}\right) \]

\newpage
Computing the posterior mean and covariance over $a$ and $b$ given by the $CO_2$ data:

\begin{figure}[h]
\centering
\begin{minipage}{.5\textwidth}
  \centering
\includegraphics[scale = 0.5]{outputs/q2/a-mean}
\caption{The Posterior Mean}
\label{fig:fig2-a-mean}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
\includegraphics[scale = 0.5]{outputs/q2/a-covariance}
\caption{The Posterior Covariance}
\label{fig:fig2-a-covariance}
\end{minipage}
\end{figure}

\subsection*{(b)}

Plotting the residuals:

\begin{figure}[h]
\centering
\includegraphics[scale=0.5]{outputs/q2/b-residuals-timeseries}
\caption{$g_{obs}(t)$}
\label{fig:b-residuals-timeseries}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[scale=0.5]{outputs/q2/b-residuals-density-estimation}
\caption{Density Estimation of Residuals vs $e(t) \sim \mathcal{N}(0, 1)$}
\label{fig:b-residuals-density-estimation}
\end{figure}

We can see that the residuals do not perfectly conform to our prior over $e(t) \sim \mathcal{N}(0, 1)$.
The density estimation shows that a mean of zero is a reasonable prior belief however the data does not seem to exhibit unit variance.
Also we know it's not iid because timeseries.

\newpage
\subsection*{(c \& d)}

We are considering the kernel:

\[k(s, t) = \theta^2 \left( \exp\left( - \frac{2\sin^2(\pi(s-t)/\tau)}{\sigma^2}\right) + \phi^2 \exp\left( - \frac{(s-t)^2}{2\eta^2}\right)\right) + \zeta^2 \delta_{s=t}\]

We can make qualitative observations this kernel by visualising the covariance (gram) matrix:

\begin{figure}[h]
\centering
\includegraphics[scale=0.5]{outputs/q2/c-gram-matrix}
\caption{Covariance Matrix}
\label{fig:c-gram-matrix}
\end{figure}

We can observe a striped pattern which indicate higher covariance at regular intervals.
This can be attributed to the sinusoidal term in the kernel and encourages sinusoidal functions.
Additionally, we can see that covariance values also decay as they are further away from the diagonal.
This can be attributed to the exponential term in the kernel, encouraging points closer in time to be more correlated and vice versa.
From our $CO_2$ data, we would want a class of functions which exhibit both of these behaviours as the data looks sinusoidal (seasonal with respect to each year) and correlations locally.

We can also visualise some samples from a Gaussian Process with the same covariance matrix and zero mean.
This verifies our observations about the covariance matrix.

\begin{figure}[h]
\centering
\includegraphics[scale=0.5]{outputs/q2/c-samples}
\caption{Samples from a zero mean GP with the provided covariance kernel}
\label{fig:c-samples}
\end{figure}

\newpage
More specifically, we can see how changing each hyper-parameter will affect the characteristics of the function.

\begin{figure}[h]
\centering
\begin{minipage}{.5\textwidth}
\includegraphics[scale = 0.35]{outputs/q2/c-parameter-samples}
\caption{Samples for different parameters}
\label{fig:fig2-c-parameter-samples}
  \centering
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
\includegraphics[scale = 0.35]{outputs/q2/c-parameter-grams}
\caption{Covariances for different parameters}
\label{fig:fig2-c-parameter-grams}
\end{minipage}
\end{figure}

\begin{itemize}
\item [$\theta$:] As $\theta$ increases, we see more pronounced periodic behavior in the sample function.
    The covariance matrix shows how increasing $\theta$ visually reveals the striped periodic component.
        This is expected because it is the parameter that adjusts the weight of the periodic component.
\item [$\sigma$:] As $\sigma$ increases, we see reduced periodic behaviour in the sample function.
    The covariance matrix shows how increasing $\sigma$ will increase covariance values in the off-diagonals.
    This is expected because it adjusts the lengthscale of the periodic portion of the kernel, which ends up dominating the function.
\item [$\phi$:] As $\phi$ increases, we see the ratio of the amplitude of the periodicity component of the sample function reduces compared to the baseline.
    The covariance matrix shows how increasing $\phi$ will start to increase the non-periodic component.
    This is expected because it adjusts the weight of the non-periodic portion of the kernel, thus the periodic component remains the same (i.e.same amplitude) but the large baseline shifts from increasing $\phi$ ends up dominating the function visually.
\item [$\eta$:] As $\eta$ increases we see smoother sample functions.
This is expected because the $\eta$ increases the lengthscale of the non-periodic component, allowing for smoother functions.
This causes the off-diagonals of the gram matrix to increase, however the periodic component is still maintained because $\eta$ doesn't affect the relative weight of the two components.
\item [$\tau$:] As $\tau$ increases, the period of the periodic function increases.
We can see this reflected in the stripes in the gram matrix getting further apart.
This makes sense because we are adjusting the period in the sinusoid function of the periodic term with $\tau$.
\item [$\zeta$:] As $\zeta$ increases, the function becomes less smooth.
    This is because the $\zeta$ parameter adjusts the weight of the $\delta_{s=t}$ parameter.
    This places stronger emphasis on the independence of each timestep, which can be seen with the reduction of relative magnitude of off-diagonals in the gram matrix.
    However, this is simply masking the periodic and squared-exponential terms as we can see with the increased magnitude of the functions as $\zeta$ increases.
\end{itemize}
\newpage
\subsection*{(e)}
Suitable values for hyper-parameters can be chosen through a combination of visual inspection and prior knowledge.
For example, it is a reasonable assumption that the $CO_2$ concentration levels have a strong yearly seasonality behaviour due to the cyclic changes in temperature, humidity, etc.
Thus we can choose $\tau=1$ to ensure functions with a period of one year to reflect this knowledge.
It can be difficult to quantitatively choose values for the other parameters as they can relate to the uncertainty exhibited in the data (i.e.the smoothness of the function).
One approach is to maximise:

\[\log P(\textbf{Y} | \textbf{X}) = -\frac{1}{2} \textbf{Y}^T (\textbf{K} + \sigma^2\textbf{I})^{-1} \textbf{Y} - \frac{1}{2} \log |\textbf{K}+ \sigma^2\textbf{I}| - \frac{n}{2}\log(2\pi)\]

the log-likelihood  of the posterior distribution with respect to the given data where $\textbf{K}$ is the gram matrix for the kernel (equation 2.30 from http://gaussianprocess.org/gpml/chapters/RW2.pdf).
We can define a loss function as the negative log-likelihood and employ gradient-based algorithms to find optimal parameters.

Comparing the hyperparameters corresponding to before and after training side by side:

\begin{figure}[h]
\centering
\begin{minipage}{.5\textwidth}
  \centering
\includegraphics[scale=0.5]{outputs/q2/f-untrained-parameters}
\caption{Untrained hyperparameters}
\label{fig:f-untrained-parameters}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
\includegraphics[scale=0.5]{outputs/q2/f-trained-parameters}
\caption{Trained Hyperparmaeters}
\label{fig:f-trained-parameters}
\end{minipage}
\end{figure}

We can analyse some of the changes in these parameters after training to gain some insights.
We can see that $\tau$ remains the same as we would expect given the yearly seasonality we have prior knowledge of.
On the other hand, the value for $\zeta$ is significantly reduced signifying that $\delta_{s=t}$ is not a very good kernel for representing the data as datapoints at different timesteps do exhibit correlations.

\newpage
\subsection*{(f)}
Extrapolating the $CO_2$ concentration levels:

\begin{figure}[h]
\centering
\begin{minipage}{.5\textwidth}
  \centering
\includegraphics[scale=0.5]{outputs/q2/f-extrapolation-untrained}
\caption{Untrained extrapolation}
\label{fig:f-extrapolation-untrained}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
\includegraphics[scale=0.5]{outputs/q2/f-extrapolation-trained}
\caption{Trained extrapolation}
\label{fig:f-extrapolation-trained}
\end{minipage}
\end{figure}

We can see that the extrapolation shows a continued increase in $CO_2$ in the future.
This follows our expectations given that the levels has been steadily increasing in the past.
Moreover, the concentration continues to exhibit yearly seasonality (for the trained extrapolation) as we would expect.
We can see that the conclusions can be quite sensitive to kernel hyperparameters when comparing the values from before and after training.
Prior to training, the extrapolated prediction is not representative of the given data, with pretty much no seasonal behaviour and very large uncertainty.
After training, we can see that the prediction is much more reasonable, and qualitatively the uncertainty bounds seem to exhibit the historical variability in the data.

\subsection*{(g)}

This procedure is not fully Bayesian because despite using a posterior estimate of our linear regression terms, we only use a point estimate when making prediction.
For a fully Bayesian approach, we should also incorporate the uncertainty of the linear regression parameters into our extrapolation/uncertainty bounds.
For our procedure, we only include the uncertainty of $g(t)$ however it can be observed in the plots that the trend is not perfectly linear so this should be reflected in the uncertainty of our extrapolation.
Another approach could be to add a linear kernel to our combined kernel function and model $f(t)$ directly with our kernel, removing the linear regression component in our procedure.
Thus our kernel extrapolation would incorporate the uncertainty of all components of our signal.

\newpage
The Python code for Bayesian Linear Regression:
\lstinputlisting[language=Python]{src/models/bayesian_linear_regression.py}

\newpage
The Python code for kernels:
\lstinputlisting[language=Python]{src/models/kernels.py}

\newpage
The Python code for Gaussian Process Regression:
\lstinputlisting[language=Python]{src/models/gaussian_process_regression.py}

\newpage
The rest of the Python code for question 2:
\lstinputlisting[language=Python]{src/solutions/q2.py}

\newpage
\section*{Question 3}

\subsection*{(a)}

The free energy is can be calculated as:
\[\mathcal{F}(q, \theta) = \left\langle \log P(\textbf{x}, \textbf{s} | \theta)\right\rangle_{q(\textbf{s})} + H[Q(\textbf{s})]\]

%Thus,
%\[\mathcal{F}(Q, \theta) = \left\langle \log P(\textbf{x}, \textbf{s} | \theta)\right\rangle_{\prod_{i=1}^{K} q_{i} (s_i)} + H\left[ \prod_{i=1}^{K} q_{i} (s_i)\right] \]

Knowing,
\[\log P(\textbf{x}, \textbf{s} | \theta) = \log P(\textbf{x} | \textbf{s}, \theta) + \log P(\textbf{s} | \theta)\]

we can write:

\[\mathcal{F}(Q, \theta) = \left\langle \log P(\textbf{x} | \textbf{s}, \theta)\right\rangle_{q(\textbf{s})} + \left\langle \log P(\textbf{s} | \theta)\right\rangle_{q(\textbf{s})} + H\left[ q(\textbf{s})\right] \]

Moreover, our mean field approximation:
\[q(\textbf{s}) = \prod_{i=1}^{K} q_{i} (s_i) \]

where $q_{i} (s_i) = \lambda_i^{s_i} (1-\lambda_i)^{(1-s_i)}$.

To compute the first term:

\[ P(\textbf{x} | \textbf{s}, \theta)  = \mathcal{N}\left( \sum_{i=1}^{K} s_i \mu_i, \sigma^2 \textbf{I} \right)}\]

substituting the appropriate terms:

\[ P(\textbf{x} | \textbf{s}, \theta)  = 2\pi^{-\frac{d}{2}} |\sigma^2 \textbf{I}|^{-\frac{1}{2}} \exp \left( -\frac{1}{2} \left(\textbf{x} - \sum_{i=1}^{K} s_i \mu_i\right)^T \frac{1}{\sigma^2} \textbf{I}  \left(\textbf{x} - \sum_{i=1}^{K} s_i \mu_i\right) \right) \]

with $d$ being the number of dimensions.

Taking the logarithm:

\[ \log P(\textbf{x} | \textbf{s}, \theta)  = -\frac{d}{2} \log (2 \pi \sigma^2)  -\frac{1}{2 \sigma^2} \left(\textbf{x}^T\textbf{x} - 2 \textbf{x}^T\sum_{i=1}^{K} s_i \mu_i   + \sum_{i=1}^{K} \sum_{i=1}^{K} s_i s_j \mu_i^T\mu_j \right) \]

The expectation distributed to the relevant terms:

\[
\left\langle \log P(\textbf{x} | \textbf{s}, \theta)\right\rangle_{q(\textbf{s})} \\
= -\frac{d}{2} \log (2 \pi \sigma^2)   -\frac{1}{2 \sigma^2} \left(\textbf{x}^T\textbf{x} - 2 \textbf{x}^T\sum_{i=1}^{K} \left\langle s_i \right\rangle_{q_{i} (s_i)} \mu_i   + \sum_{i=1}^{K} \sum_{j=1}^{K} \left\langle s_i s_j \right\rangle_{q_{i} (s_i) q_{j} (s_j)} \mu_i^T\mu_j \right)\]

Evaluating the expectations:

\[
\left\langle \log P(\textbf{x} | \textbf{s}, \theta)\right\rangle_{q(\textbf{s})} \\
= -\frac{d}{2} \log (2 \pi \sigma^2)  -\frac{1}{2 \sigma^2} \left(\textbf{x}^T\textbf{x} - 2 \textbf{x}^T\sum_{i=1}^{K}  \lambda_i  \mu_i   + \sum_{i=1}^{K} \sum_{j=1, j\neq i}^{K}  \lambda_i \lambda_j \mu_i^T\mu_j + \sum_{i=1}^{K}  \lambda_i \mu_i^T\mu_i \right)\]

where $\left\langle s_i s_i \right\rangle_{q_{i} (s_i)} = \left\langle s_i \right\rangle_{q_{i} (s_i)}$ because $s_i \in \{0, 1\}$.

To compute the second term:
\[ P(\textbf{s} | \theta) = \prod_{i=1}^{K}\pi_i^{s_i} (1-\pi_i)^{(1-s_i)}\]

Taking the logarithm:

\[ \log P(\textbf{s} | \theta) = \sum_{i=1}^{K}s_i \log \pi_i + (1-s_i) \log (1-\pi_i)}\]

The expectation distributed to the relevant terms:

\[ \left\langle \log P(\textbf{s} | \theta)\right\rangle_{q(\textbf{s})}= \sum_{i=1}^{K} \left\langle s_i \right\rangle_{q_{i} (s_i)} \log \pi_i + (1-\left\langle s_i\right\rangle_{q_{i} (s_i)}) \log (1-\pi_i)}\]

Evaluating the expectations:

\[ \left\langle \log P(\textbf{s} | \theta)\right\rangle_{q(\textbf{s})}= \sum_{i=1}^{K} \lambda_i \log \pi_i + (1-\lambda_i) \log (1-\pi_i)}\]

To compute the third term, we use the mean field factorisation:

\[H\left[ q(\textbf{s})\right] = \sum_{i=1}^K H\left[ q_{i} (s_i)\right] \]

Thus,

\[H\left[ q(\textbf{s})\right] = - \sum_{i=1}^K \sum_{s_i \in \{0, 1\}} q_{i} (s_i) \log q_{i} (s_i) \]

Substituting the appropriate values:

\[H\left[ q(\textbf{s})\right] = - \sum_{i=1}^K \lambda_i \log \lambda_i + (1-\lambda_i) \log (1-\lambda_i) \]

Combining, we have our free energy expression:

\[
\begin{array}{l}
\tab \mathcal{F}(q, \theta) = \\
\tab \tab \frac{-d}{2} \log (2 \pi \sigma^2)   -\frac{1}{2 \sigma^2} \left(\textbf{x}^T\textbf{x} - 2 \textbf{x}^T\sum_{i=1}^{K}  \lambda_i  \mu_i   + \sum_{i=1}^{K} \sum_{j=1, j\neq i}^{K}  \lambda_i \lambda_j \mu_i^T\mu_j + \sum_{i=1}^{K}  \lambda_i \mu_i^T\mu_i \right) \\
\tab \tab + \sum_{i=1}^{K} \lambda_i \log \pi_i + (1-\lambda_i) \log (1-\pi_i)}\\
\tab \tab - \sum_{i=1}^K \lambda_i \log \lambda_i + (1-\lambda_i) \log (1-\lambda_i)
\end{array}
\]

To derive the partial update for $q_i(s_i)$ we take the variational derivative of the Lagrangian, enforcing the normalisation of $q_i$:

\[\frac{\partial}{\partial q_i}\left( \mathcal{F}(q, \theta) + \lambda^{LG} \int q_i -1)\right) = \left\langle \log P(\textbf{x}, \textbf{s} | \theta)\right\rangle_{\prod_{j\neq i} q_j(s_j)} - \log q_i(s_i) - 1 + \lambda^{LG}\]

where $\lambda^{LG}$ is the Lagrange multiplier.

Setting this to zero we can solve for the $\lambda_i$ that maximises the free energy:

\[\log q_i(s_i) = \left\langle \log P(\textbf{x}, \textbf{s} | \theta)\right\rangle_{\prod_{j\neq i} q_j(s_j)} - 1 + \lambda^{LG}\]

Similar to our free energy derivation:

\[\left\langle \log P(\textbf{x} |  \textbf{s}, \theta)\right\rangle_{\prod_{j\neq i} q_j(s_j)} \propto  -\frac{1}{2 \sigma^2} \left(\textbf{x}^T\textbf{x} - 2 \textbf{x}^T\sum_{k=1}^{K} \left\langle s_k \right\rangle_{\prod_{j\neq i} q_j(s_j)} \mu_i   + \sum_{k=1}^{K} \sum_{j=1}^{K} \left\langle s_k s_j \right\rangle_{\prod_{j\neq i} q_j(s_j)} \right)\]

and

\[ \left\langle \log P(\textbf{s} | \theta)\right\rangle_{\prod_{j\neq i} q_j(s_j)} = \sum_{k=1}^{K} \left\langle s_k \right\rangle_{\prod_{j\neq i} q_j(s_j)} \log \pi_k + (1-\left\langle s_k\right\rangle_{\prod_{j\neq i} q_j(s_j)}) \log (1-\pi_k)}\]

We can write:

\[\log q_i(s_i) \propto  \log P(\textbf{x} |  \textbf{s}, \theta)\right\rangle_{\prod_{j\neq i} q_j(s_j)} + \left\langle \log P(\textbf{s} | \theta)\right\rangle_{\prod_{j\neq i} q_j(s_j)} \]

Substituting the relevant terms:

\[\log q_i(s_i) \propto  -\frac{1}{2 \sigma^2} \left( - 2 s_i \textbf{x}^T \mu_i   + s_i s_i  \mu_i^T\mu_i +  2 \sum_{j=1, j\neq i}^{K}  s_i \lambda_j  \mu_i^T\mu_j  \right) + s_i \log \pi_i + (1- s_i ) \log (1-\pi_i)}
\]

Knowing $\log q_i(s_i) = s_i \log \lambda_i + (1-s_i) \log (1-\lambda_i)$:

\[\log q_i(s_i) \propto s_i \log \frac{\lambda_i}{1-\lambda_i}\]

Thus,

\[s_i \log \frac{\lambda_i}{1-\lambda_i} \propto  -\frac{1}{2 \sigma^2} \left( - 2 s_i \textbf{x}^T \mu_i   + s_i s_i  \mu_i^T\mu_i +  2 \sum_{j=1, j\neq i}^{K}  s_i \lambda_j  \mu_i^T\mu_j  \right) + s_i \log \frac{\pi_i}{1-\pi_i}}\]

Also, because $s_i \in \{0, 1\}}$ we know that $s_i^2 = s_i$:

\[s_i \log \frac{\lambda_i}{1-\lambda_i} \propto  -\frac{1}{2 \sigma^2} \left( - 2 s_i \textbf{x}^T \mu_i   +  s_i  \mu_i^T\mu_i +  2 \sum_{j=1, j\neq i}^{K}  s_i \lambda_j  \mu_i^T\mu_j  \right) + s_i \log \frac{\pi_i}{1-\pi_i}}\]

Because we have only kept terms with $s_i$, this is an equality:

\[s_i \log \frac{\lambda_i}{1-\lambda_i} =  \frac{  s_i \mu_i^T}{2 \sigma^2} \left( 2\textbf{x} - \mu_i -  2 \sum_{j=1, j\neq i}^{K}   \lambda_j  \mu_j \right) + s_i \log \frac{\pi_i}{1-\pi_i}}\]

Solving for $\lambda_i$:

\[ \lambda_i =  \frac{1}{1+\exp\left[ - \left(\frac{  \mu_i^T}{\sigma^2} \left( \textbf{x} -  \frac{\mu_i}{2} -  \sum_{j=1, j\neq i}^{K}   \lambda_j  \mu_j \right) + \log \frac{\pi_i}{1-\pi_i}}\right) \right]}\]

we have our partial update.

\newpage

\subsection*{(b)}

The provided derivations for the M step of the mean parameter $\mu$:

\[\mu = \left( \left\langle \textbf{s}\textbf{s}^T \right\rangle_{q(\textbf{s})} \right)^{-1}\left\langle \textbf{s}\right\rangle_{q(\textbf{s})} \textbf{x}\]

where $\mu \in \mathbb{R}^{K \times D}$, \textbf{s} \in \mathbb{R}^{K \times N}$, and $\textbf{x} \in \mathbb{R}^{N \times D}$.

This mimics the least squares solution:

\[\hat{\beta} = (\textbf{X}\textbf{X}^T) \textbf{X} \textbf{Y}\]

for the linear regression problem $\textbf{Y} = \textbf{X}^T \beta$ where $\beta$ corresponds to the mean parameters $\mu$, the design matrix $\textbf{X}$ corresponds to the input $\textbf{s}$ and the response $Y$ corresponds to the image pixels denoted $\textbf{x}$.
This makes sense because our resulting images $\textbf{x}$ are modeled as linear combinations of features $\mu$, weighted by $\textbf{s}$.

\subsection*{(c)}

The computational complexity of the implemented M step function can be broken down for each parameter:

\begin{itemize}
\item [$\mu$:]\begin{itemize}
\item The inversion $\text{ESS}^{-1}$ where $\text{ESS} \in \mathbb{R}^{K \times K}$ is $\mathcal{O}(K^3)$
\item The dot product $\text{ESS}^{-1} \text{ES}^T$ where $\text{ESS}^{-1} \in \mathbb{R}^{K \times K}$ and $\text{ES} \in \mathbb{R}^{N \times K}$ is  $\mathcal{O}(K^2N)$
\item The dot product $(\text{ESS}^{-1} \text{ES}^T)\textbf{x}$ where $(\text{ESS}^{-1} \text{ES}^T)\ \in \mathbb{R}^{K \times N}$ and $\textbf{x} \in \mathbb{R}^{N \times D}$ is  $\mathcal{O}(KND )$
\end{itemize}

\item [$\sigma$:] \begin{itemize}
\item The dot product $(\textbf{x}^{T} \textbf{x})$ where $\textbf{x} \in \mathbb{R}^{N \times D}$ is  $\mathcal{O}(D^2N)$
\item The dot product $\mu^{T} \mu$ where $\mu \in \mathbb{R}^{D \times K}$ is  $\mathcal{O}(K^2D)$
\item The dot product $(\mu^{T} \mu)\text{ESS}$ where $\mu^{T} \mu \in \mathbb{R}^{K \times K}$ and $\text{ESS} \in \mathbb{R}^{K \times K}$ is $\mathcal{O}(K^3)$
\end{itemize}
\item [$\pi$:] \begin{itemize}
    \item The mean operation for $\text{ES} \in \mathbb{R}^{N \times K}$ along the first dimension is $\mathcal{O}(NK)$
    \end{itemize}
\end{itemize}

Thus, the computational complexity of the M step is $\mathcal{O}(K^3+K^2N + KND + D^2N + K^2D)$ where we do not assume that any of $N$, $K$, or $D$ is large compared to the others.
\newpage
\subsection*{(d)}

\begin{figure}[h]
\centering
\includegraphics[scale=0.4]{data/images}
\caption{Images generated by randomly combined features with noise}
\label{fig:3d-images}
\end{figure}

\subsection*{(e)}

We can plot the free energy to make sure it increases each iteration:

\subsection*{(f)}

\subsection*{(g)}

\newpage
The Python code for the binary latent factor model:
\lstinputlisting[language=Python]{src/models/binary_latent_factor_model.py}

\newpage
The Python code for mean field learning:
\lstinputlisting[language=Python]{src/models/mean_field_learning.py}

\newpage
The rest of the Python code for question 3:
\lstinputlisting[language=Python]{src/solutions/q3.py}
\newpage
\section*{Question 4}

We begin with the log joint:

\[\log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) = \log P(\textbf{x}|  \textbf{s}, \textbf{A}, \Psi,  \eta)
+ \log P( \textbf{s}| \pi, \eta) + \log P( \pi| \eta) + \log P( \textbf{A}| \eta) + \log P( \Psi| \eta)
\]

where $\eta$ is a collection of all hyperparameters.

We know:

\[P(\textbf{x}|  \textbf{s}, \textbf{A}, \Psi,  \eta) = \frac{1}{(2\pi)^{\frac{D}{2}} |\Psi|^{\frac{1}{2}}} \exp\left(-\frac{1}{2}(\textbf{x}-\textbf{As})^T \Psi^{-1} (\textbf{x}-\textbf{As})\right)\]

\[P( \textbf{s}| \pi, \eta) = \prod_{k=1}^{K} \pi_k^{s_i} (1-\pi_k)^{1-s_k}\]

\[P( \pi| \eta) = \prod_{k=1}^{K} \frac{\pi_k^{\alpha - 1}(1-\pi_k)^{\beta-1}}{B(\alpha, \beta)}\]

For $\textbf{A}$ we choose a factorised conjugate prior:

\[P( \textbf{A}| \eta) = \prod_{k=1}^{K}P( \textbf{A}_{: k}| \eta)\]

where $\textbf{A}_{: k} \in \mathbb{R}^{D \times 1}$ is the $k^{th}$ column of $\textbf{A}$. For each column we choose:

\[P( \textbf{A}_{: k}| \eta) = \mathcal{N}(\textbf{A}_{: k}| \mu_{\textbf{A}_{: k}}, \Sigma_{\textbf{A}_{: k}}) =\frac{1}{(2\pi)^{\frac{D}{2}} |\Sigma_{\textbf{A}_{: k}}|^{\frac{1}{2}}} \exp\left(-\frac{1}{2}(\textbf{A}_{: k} - \mu_{\textbf{A}_{: k}})^T \Sigma_{\textbf{A}_{: k}}^{-1}(\textbf{A}_d - \mu_{\textbf{A}_d})\right)\]

a Gaussian prior with diagonal covariance $\Sigma_{\textbf{A}_{: k}} = \alpha_k^{2}\textbf{I}$ and mean zero, so we can simplify:

\[P( \textbf{A}_{: k}| \alpha_k) = (2\pi\alpha_k^{2})^{\frac{-D}{2}} \exp\left(-\frac{\textbf{A}_{: k}^T \textbf{A}_{: k}}{2\alpha_k^{2}} \right)\]

For $\Psi}$ we choose a conjugate prior:

\[P( \Psi| \eta) = \prod_{d=1}^D InvGamma(\Psi_{dd}| a, b) = \prod_{d=1}^D \frac{b^{a}}{\Gamma(a)}\Psi_{dd}^{-a-1}\exp(-\frac{b}{\Psi_{dd}})\]

a product of inverse gamma distributions on $\Psi$ where we assume $\Psi$ is a diagonal matrix.

Combining, we have our expression:

\[
\begin{array}{l}
\log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) = -\frac{D}{2} \log (2 \pi) - \frac{1}{2} \sum_{d=1}^{D}\log\Psi_{dd}   -\frac{1}{2}(\textbf{x}-\textbf{As})^T \Psi^{-1} (\textbf{x}-\textbf{As}) \\
\tab \tab \tab \tab \tab + \sum_{k=1}^{K} s_k \log \pi_k + (1-s_k) \log (1-\pi_k)}\\
\tab \tab \tab \tab \tab + \sum_{i=1}^K (\alpha-1) \log \pi_k + (\beta - 1) \log (1-\pi_k) - \log B(\alpha, \beta)\\
\tab \tab \tab \tab \tab + \sum_{i=1}^K -\frac{D}{2} \log (2 \pi\alpha_k^2)  -\frac{\textbf{A}_{: k}^T \textbf{A}_{: k}}{2\alpha_k^{2}}\\
\tab \tab \tab \tab \tab + \sum_{d=1}^D a \log b + (-a - 1) \log\Psi_{dd} - \frac{b}{\Psi_{dd}} - \log \Gamma(a)\\
\end{array}
\]

%Simplifying:
%
%\[
%\begin{array}{l}
%\log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) = -\frac{D}{2} \log (2 \pi)  -\frac{1}{2}(\textbf{x}-\textbf{As})^T \Psi^{-1} (\textbf{x}-\textbf{As}) \\
%\tab \tab \tab \tab \tab + \sum_{i=1}^K (s_k+\alpha-1) \log \pi_k + (\beta - s_k) \log (1-\pi_k) - \log B(\alpha, \beta)\\
%\tab \tab \tab \tab \tab -\frac{DK}{2} \log (2 \pi) - \frac{1}{2} \log(|\Sigma_{\textbf{A}}|)   -\frac{1}{2}\textbf{A}^T \Sigma_{\textbf{A}}^{-1}\textbf{A}\\
%\tab \tab \tab \tab \tab + \sum_{d=1}^D a_d \log b_d + (-a_d - \frac{3}{2}) \log\Psi_{dd} - \frac{b_d}{\Psi_{dd}} - \log \Gamma(a_d)\\
%\end{array}
%\]

For the Variational Bayes expectation step, we minimise $\textbf{KL}[q_{\texbf{s}}(\textbf{s} | \text{everything else}) \| P(\textbf{s} | \text{everything else})]$ by setting:

\[q_{\texbf{s}}(\textbf{s}) \propto \exp \left\langle \log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) \right\rangle_{q(\theta)}\]

where $\theta$ denotes the parameters $\pi$, $\textbf{A}$, $\Psi$, $\eta$.

Substituting the relevant terms:

\[q_{\texbf{s}}(\textbf{s}) \propto \exp \left\langle -\frac{1}{2}(\textbf{x}-\textbf{As})^T \Psi^{-1} (\textbf{x}-\textbf{As}) + \sum_{k=1}^{K} s_k \log \pi_k + (1-s_k) \log (1-\pi_k)}\right\rangle_{q(\theta)}\]

Simplifying:

\[q_{\texbf{s}}(\textbf{s}) \propto \exp \left\langle -\frac{1}{2}\left(\textbf{s}^T\textbf{A}^T\Psi^{-1}\textbf{As} - 2\textbf{s}^T\left(\textbf{A}^T\Psi^{-1} \textbf{x}  + 2\log\frac{\pi}{1-\pi}\right)\right)\right\rangle_{q(\theta)}\]

By inspection, we can see:

\[q_{\texbf{s}}(\textbf{s}) \propto \mathcal{N}(s | \mu_\textbf{s}^*, \Sigma_\textbf{s}^*)\]

where

\[\Sigma_\textbf{s}^* = \left\langle \left(\textbf{A}^T\Psi^{-1}\textbf{A}\right)^{-1}\right\rangle_{q(\theta)}\]

and

\[\mu_\textbf{s}^* = \left\langle \left(\textbf{A}^T\Psi^{-1}\textbf{A}\right)^{-1}\left(\textbf{A}^T\Psi^{-1} \textbf{x}  + 2\log\frac{\pi}{1-\pi}\right)\right\rangle_{q(\theta)}\]

the E step updates.

For the Variational Bayes maximisation step, we set:

\[q_{\theta}(\theta) \propto P(\theta) \exp \left\langle \log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) \right\rangle_{q(\textbf{s})}\]

assuming the factorisation:

\[q_{\theta}(\theta) = q_{\pi}(\pi)q_{\Psi}(\Psi)q_{\textbf{A}}(\textbf{A})\]

we can calculate each factor independently.

For $q_{\pi}(\pi)$:

\[q_{\pi}(\pi) \propto P(\pi)\exp \left\langle \log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \pi}(\theta)}\]

Substituting the appropriate terms:

\[q_{\pi}(\pi) \propto \left(\prod_{k=1}^{K} \frac{\pi_k^{\alpha - 1}(1-\pi_k)^{\beta-1}}{B(\alpha, \beta)} \right)\exp \left\langle \sum_{i=1}^K s_k \log \pi_k + (1 - s_k) \log (1-\pi_k) \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \pi}(\theta)}\]

We see:

\[q_{\pi}(\pi) \propto \prod_{k=1}^{K} \frac{\pi_k^{\alpha  + \left\langle s_k\right\rangle_{q_{s_k}} - 1}(1-\pi_k)^{\beta-\left\langle s_k\right\rangle_{q_{s_k}}}}{B(\alpha, \beta)}\]

\[q_{\pi}(\pi) = \prod_{k=1}^{K}  Beta(\alpha + \left\langle s_k\right\rangle_{q_{s_k}} , \beta+(1-\left\langle s_k\right\rangle_{q_{s_k}}}))\]

%Substituting the appropriate terms:
%
%\[q_{\pi}(\pi) \propto \left(\prod_{k=1}^{K} \frac{\pi_k^{\alpha - 1}(1-\pi_k)^{\beta-1}}{B(\alpha, \beta)} \right)\exp \left\langle \sum_{i=1}^K (s_k+\alpha-1) \log \pi_k + (\beta - s_k) \log (1-\pi_k) \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \pi}(\theta)}\]
%
%We see:
%
%\[q_{\pi}(\pi) \propto \prod_{k=1}^{K} \frac{\pi_k^{2\alpha - 2 + \left\langle s_k\right\rangle_{q_{s_k}}}(1-\pi_k)^{2\beta-\left\langle s_k\right\rangle_{q_{s_k}}}}{B(\alpha, \beta)}\]
%
%\[q_{\pi}(\pi) = \prod_{k=1}^{K}  Beta(2\alpha -2 + \left\langle s_k\right\rangle_{q_{s_k}} , 2\beta+1-\left\langle s_k\right\rangle_{q_{s_k}}})\]

For $q_{\Psi}(\Psi)$:

\[q_{\Psi}(\Psi) \propto P(\Psi)\exp \left\langle \log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \Psi}(\theta)}\]

Substituting the appropriate terms:

\[
q_{\Psi}(\Psi) \propto \left( \prod_{d=1}^D \frac{b^{a}}{\Gamma(a)}\Psi_{dd}^{-a-1}\exp(-\frac{b}{\Psi_{dd}}) \right)\exp \left\langle  - \frac{1}{2} \sum_{d=1}^{D}\log\Psi_{dd}   -\frac{1}{2}(\textbf{x}-\textbf{As})^T \Psi^{-1} (\textbf{x}-\textbf{As}) \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \Psi(\theta)}}
\]

We see:

\[
q_{\Psi}(\Psi) \propto  \prod_{d=1}^D \frac{b^{a}}{\Gamma(a)}\Psi_{dd}^{-(a+\frac{1}{2})-1}\exp\left(-\frac{b+\frac{1}{2} \left\langle (x_d-\textbf{A}_{d :}\textbf{s})^2 \right\rangle_{q_{\textbf{s}}(\textbf{s})q_{\textbf{A}_{d :}}(\textbf{A}_{d :})}}{\Psi_{dd}}\right) }
\]

where $\textbf{A}_{d :} \in \mathbb{R}^{1 \times K}$ is the $d^{th}$ row of $\textbf{A}$.

Thus,

\[
q_{\Psi}(\Psi) = \prod_{d=1}^D  InvGamma\left(\Psi_{dd} \left| a+\frac{1}{2}, b+\frac{1}{2} \left\langle (x_d-\textbf{A}_{d :}\textbf{s})^2 \right\rangle_{q_{\textbf{s}}(\textbf{s})q_{\textbf{A}_{d :}}(\textbf{A}_{d :})}\right)
\]

For $q_{\textbf{A}_{: k}}(\textbf{A}_{: k})$:

\[q_{\textbf{A}_{: k}}(\textbf{A}_{: k})  \propto P(\textbf{A}_{: k})\exp \left\langle \log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \textbf{A}_{: k}}(\theta)}\]

Substituting the appropriate terms:

\[q_{\textbf{A}_{: k}}(\textbf{A}_{: k}) \propto \exp\left(-\frac{\textbf{A}_{: k}^T \textbf{A}_{: k}}{2\alpha_k^{2}} \right) \exp \left\langle -\frac{1}{2}(\textbf{x}-\textbf{A}_{: k} s_k)^T \Psi^{-1} (\textbf{x}-\textbf{A}_{: k} s_k) \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \textbf{A}_{: k}}(\theta)}\]

\[q_{\textbf{A}_{: k}}(\textbf{A}_{: k}) \propto \exp\left(-\frac{\textbf{A}_{: k}^T \textbf{A}_{: k}}{2\alpha_k^{2}} \right) \exp \left\langle -\frac{1}{2} \textbf{A}_{: k}^T\frac{1 }{ s_k^2 \Psi_{dd}}\textbf{A}_{: k}  - 2 \frac{s_k}{\Psi_{dd}} \textbf{x}^T\textbf{A}_{: k} \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \textbf{A}_{: k}}(\theta)}\]

We see that

\[q_{\textbf{A}_{: k}}(\textbf{A}_{: k}) = \mathcal{N}\left(\mu_{\textbf{A}_{: k}}, \Sigma_{\textbf{A}_{: k}}\right)}\]

where:

\[ \Sigma_{\textbf{A}_{: k}} = \left(\frac{1}{\alpha_k^2} + \frac{1}{s_k^2\Psi_{dd}}\right)^{-1} \textbf{I}\]

and

\[\mu_{\textbf{A}_{: k}} = \Sigma_{\textbf{A}_{: k}}^{-1} \frac{s_k}{\Psi_{dd}}\textbf{x}\]

By optimising with respect to the the distributions $\Psi$ and $\alpha$ in turn causes some $\alpha_i^2$ to diverge, the number of remaining $\alpha_i^2$ provide our determination for the value of $K$. This is automatic relevance determination through factor analysis.

\newpage
\section*{Question 5}

\subsection*{(a)}

The log-joint probability for a single observation-source pair:

\[\log p(\textbf{s}, \textbf{x}) = \log p(\textbf{s}) + \logp(\textbf{x}|\textbf{s})\]

Knowing $p(\textbf{s}) = \prod_{i=1}^{K}p(s_i| \pi_i)$ and $p(\textbf{x}|\textbf{s}) = \mathcal{N}(\sum_{i=1}^{K} s_i \mu_i, \sigma^2 \textbf{I})$:

\[\log p(\textbf{s}, \textbf{x})  \propto \frac{-1}{2}\left( \textbf{x} - \sum_{i=1}^{K}s_i \mu_i\right)^T \frac{1}{\sigma^2} \textbf{I} \left( \textbf{x} - \sum_{i=1}^{K} s_i \mu_i\right) + \sum_{i=1}^{K} \left(s_i \log\pi_i + (1-s_i)\log(1-\pi_i)\right)\]

Expanding:

\[\log p(\textbf{s}, \textbf{x})  \propto \frac{-1}{2\sigma^2} \left( \textbf{x}^T\textbf{x} - 2\textbf{x}^T\sum_{i=1}^{K}s_i \mu_i + \sum_{i=1}^{K}\sum_{j=1}^{K}s_i s_j \mu_i^T \mu_j\right) + \sum_{i=1}^{K} \left(s_i \log\pi_i + (1-s_i)\log(1-\pi_i)\right)\]

Collecting terms pertaining to $s_i$:

\[\log p(\textbf{s}, \textbf{x})  =    \sum_{i=1}^{K} \left(\left(\frac{\textbf{x}^T \mu_i}{\sigma^2} +\log\frac{\pi_i}{1-\pi_i} \right) s_i\right) + \sum_{i=1}^{K}\sum_{j=1}^{K}\left( \frac{ - \mu_i^T \mu_j}{2\sigma^2} s_i s_j \right) + C\]

where $C$ are all other terms without $s_i$.

Knowing that $s_i^2= s_i$:

\[\log p(\textbf{s}, \textbf{x})  =    \sum_{i=1}^{K} \left(\left(\frac{\textbf{x}^T \mu_i}{\sigma^2} +\log\frac{\pi_i}{1-\pi_i} - \frac{\mu_i^T \mu_i}{2\sigma^2} \right) s_i\right) + \sum_{i=1}^{K}\sum_{j=1}^{i-1}\left( \frac{ - \mu_i^T \mu_j}{\sigma^2} s_i s_j \right) + C\]

Thus:

\[\log p(\textbf{s}, \textbf{x})  =    \sum_{i=1}^{K} \log f_i(s_i) + \sum_{i=1}^{K}\sum_{j=1}^{i-1}\log g_{ij}(s_i, s_j)\right) \]

where the factors are defined:

\[\log f_i(s_i) = \left(\frac{\textbf{x}^T \mu_i}{\sigma^2} +\log\frac{\pi_i}{1-\pi_i} - \frac{\mu_i^T \mu_i}{2\sigma^2} \right) s_i\]

and

\[\log g_{ij}(s_i, s_j) = \frac{- \mu_i^T \mu_j}{\sigma^2} s_i s_j\]

as required.

The Boltzmann Machine can be defined:

\[P(\textbf{s}| \textbf{W}, \textbf{b}) = \frac{1}{Z} \exp\left( \sum_{i=1}^{K}\sum_{j=1}^{i-1}W_{ij}s_i s_j - \sum_{i=1}^{K} b_i s_i\right)}\]

where $s_i \in \{0, 1\}$, the same as our source variables.

From our factorisation, we can see that $p(\textbf{s}, \textbf{x})$ is a Boltzmann Machine with:

\[W_{ij} = \frac{- \mu_i^T \mu_j}{\sigma^2}\]

and

\[b_i = -\left(\frac{\textbf{x}^T \mu_i}{\sigma^2} +\log\frac{\pi_i}{1-\pi_i} - \frac{\mu_i^T \mu_i}{2\sigma^2}\right)\]

and

\[\log Z = -C \]

\subsection*{(b)}

For $f_i(s_i)$, we will choose a Bernoulli approximation:

\[\tilde{f}_i(s_i) = \lambda_i^{s_i} + (1-\lambda_i)^{1-s_i}\]

Thus,

\[\log \tilde{f}_i(s_i) \propto \log \left(\frac{\lambda_i}{1-\lambda_i} \right)s_i\]

For $g_{ij}(s_i, s_j)$, we will choose a product of Bernoulli's approximation:

\[\tilde{g}_{ij}(s_i, s_j) = \tilde{g}_{ij, \neg s_j}(s_i)\tilde{g}_{ij, \neg s_i}(s_j)\]

where

\[\tilde{g}_{ij, \neg s_j}(s_i) = (\theta_{ji})^{s_i} + (1-\theta_{ji})^{1-s_i}\]

and

\[\tilde{g}_{ij, \neg s_i}(s_j) = (\theta_{ij})^{s_j} + (1-\theta_{ij})^{1-s_j}\]

Thus,

\[\log \tilde{g}_{ij}(s_i, s_j) \propto \log \left(\frac{\theta_{ji}}{1-\theta_{ji}} \right) s_i + \log \left(\frac{\theta_{ij}}{1-\theta_{ij}} \right) s_j\]

we can define $\xi_{ji} = \log \left(\frac{\theta_{ji}}{1-\theta_{ji}} \right)$ and $\xi_{ij} = \log \left(\frac{\theta_{ij}}{1-\theta_{ij}} \right)$:

\[\log \tilde{g}_{ij}(s_i, s_j) \propto \xi_{ji} s_i + \xi_{ij} s_j\]

To derive the a message passing scheme, we first define the incoming message to node $i$ from the singleton factor:

\[\mathcal{M}_{i}(s_i) = \tilde{f}_i(s_i)\]

and the message incoming message to node $i$ from node $j$:

%\[\mathcal{M}_{j \rightarrow i} = \sum_{\substack{s_m \in \{0, 1\} \\ m \in \{1, ..., j-1, j+1,... K\}}} \tilde{f}_j(s_j) \prod_{k\in ne(j), k\neq i}^{K}} g_{kj}(s_k, s_j)\]

%\[\mathcal{M}_{j \rightarrow i}(s_i) = \sum_{s_1 \in \{0, 1\} } \cdots \sum_{s_{i-1} \in \{0, 1\} } \sum_{s_{i+1} \in \{0, 1\} } \cdots \sum_{s_1 \in \{0, 1\} } \tilde{f}_j(s_j) \tilde{g}_{ji}(s_j, s_i) \prod_{k\in ne(j), k\neq i}^{K}} \tilde{g}_{kj}(s_k, s_j)\]

\[\mathcal{M}_{j \rightarrow i}(s_i) = \sum_{s_1 \in \{0, 1\} } \cdots \sum_{s_{i-1} \in \{0, 1\} } \sum_{s_{i+1} \in \{0, 1\} } \cdots \sum_{s_1 \in \{0, 1\} } \tilde{f}_j(s_j) \tilde{g}_{ji}(s_j, s_i) \prod_{k\in ne(j), k\neq i}^{K}} \mathcal{M}_{k \rightarrow j}(s_j)\]

where $ne(j)$ are indices of neighbouring nodes of node $j$.

Because $\tilde{g}_{ji}(s_j, s_i)$ is a product:

\[\mathcal{M}_{j \rightarrow i}(s_i) = \tilde{g}_{ji, \neg s_j}(s_i) \sum_{s_1 \in \{0, 1\} } \cdots \sum_{s_{i-1} \in \{0, 1\} } \sum_{s_{i+1} \in \{0, 1\} } \cdots \sum_{s_1 \in \{0, 1\} } \tilde{f}_j(s_j) \tilde{g}_{ji, \neg s_i}(s_j) \prod_{k\in ne(j), k\neq i}^{K}} \mathcal{M}_{k \rightarrow j}(s_j)\]

Simplifying:

\[\mathcal{M}_{j \rightarrow i}(s_i) = \tilde{g}_{ji, \neg s_j}(s_i)\]

and,

\[\mathcal{M}_{j \rightarrow i}(s_i) \propto \exp\left(\xi_{ji} s_i\right)\]

%\[\mathcal{M}_{j \rightarrow i}(s_i) = \tilde{g}_{ji, \neg s_j}(s_i)\]

%\[\mathcal{M}_{j \rightarrow i} = \tilde{f}_j(s_j) \prod_{k\in ne(j), k\neq i}^{K}}\tilde{g}_{kj, \neg s_k}(s_j)  \sum_{\substack{s_m \in \{0, 1\} \\ m \in \{1, ..., j-1, j+1,... K\}}} \prod_{k\in ne(j), k\neq i}^{K}} \tilde{g}_{kj, \neg s_j}(s_k)\]

%\[\mathcal{M}_{j \rightarrow i}(s_j) = \tilde{f}_j(s_j) \prod_{k\in ne(j), k\neq i}^{K}}\tilde{g}_{kj, \neg s_k}(s_j) \prod_{k\in ne(j), k\neq i}^{K}} \left(\sum_{s_k \in \{0, 1\} } \tilde{g}_{kj, \neg s_j}(s_k) \right)\]
%
%\[\mathcal{M}_{j \rightarrow i}(s_j) = \tilde{f}_j(s_j) \prod_{k\in ne(j), k\neq i}^{K}}\tilde{g}_{kj, \neg s_k}(s_j)\]

Thus, the cavity distributions are:

\[q_{\neg \tilde{f}_i(s_i)}(s_i) =
\prod_{j\in ne(i)}^{K} \mathcal{M}_{j \rightarrow i}(s_i)
\]

and

\[q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j) = \left( \mathcal{M}_{i}(s_i) \prod_{k\in ne(i), k\neq j}^{K} \mathcal{M}_{k \rightarrow i}(s_i)
\right)\left( \mathcal{M}_{j}(s_j) \prod_{k\in ne(j), k\neq i}^{K} \mathcal{M}_{k \rightarrow j}(s_j)
\right)
\]

For $\tilde{f}_{i}(s_{i})$, we do not need to make an approximation step.
This is because we are minimising:

\[\tilde{f}_{i}(s_{i}) = \arg \min_{\tilde{f}_{i}(s_{i})} \textbf{KL} \left[ f_{i}(s_{i}) q_{\neg \tilde{f}_i(s_i)}(s_i) \| \tilde{f}_{i}(s_{i}) q_{\neg \tilde{f}_i(s_i)}(s_i) \right]\]

We know that the factor $\log f_i(s_i)$ is a Bernoulli of the form $b_i s_i$. Because our approximation for this site is also Bernoulli, we can simply solve for $\lambda_i$ in $\log \tilde{f}_{i}(s_{i})$:

\[\log \tilde{f}_i(s_i) = \log f_{i}(s_{i})\]

\[\log \left(\frac{\lambda_i}{1-\lambda_i} \right)s_i = b_i s_i\]

\[\lambda_i = \frac{1}{1+\exp(-b_i)}\]

On the other hand, for $\tilde{g}_{ij}(s_i, s_j)$, we will approximate with:

\[\tilde{g}_{ij}(s_i, s_j) = \arg \min_{\tilde{g}_{ij}(s_i, s_j)} \textbf{KL} \left[ g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j) \left\| \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j) \right]\]

We can define natural parameters $\eta_{i, \neg s_j}$ and $\eta_{j, \neg s_i}$ for $q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j)$ such that:

\[\mathcal{M}_{i}(s_i) \prod_{k\in ne(i), k\neq j}^{K} \mathcal{M}_{k \rightarrow i}(s_i) \propto \exp(\eta_{i, \neg s_j} s_i)\]

\[ \mathcal{M}_{j}(s_j) \prod_{k\in ne(j), k\neq j}^{K} \mathcal{M}_{k \rightarrow j}(s_j) \propto \exp(\eta_{j, \neg s_i} s_j)\]

where:

\[\eta_{i, \neg s_j} =  \log \left(\frac{\lambda_i}{1-\lambda_i}\right) + \sum_{k\in ne(i), k\neq j}^{K} \log\left(\frac{\theta_{ki}}{1-\theta_{ki}} \right)\]

Knowing $b_i=\log \left(\frac{\lambda_i}{1-\lambda_i}\right)$ and $\xi_{ki}=\log\left(\frac{\theta_{ki}}{1-\theta_{ki}} \right)$:

\[\eta_{i, \neg s_j} =  b_i + \sum_{k\in ne(i), k\neq j}^{K} \xi_{ki}\]

and

\[\eta_{j, \neg s_i} =  b_j + \sum_{k\in ne(j), k\neq i}^{K} \xi_{kj}\]

Note that $\tilde{g}_{ij}(s_i, s_j)$ was chosen as the product of two Bernoulli distributions, updates to this site approximation involves updating the  parameters $\xi_{ij}$ and $\xi_{ji}$, for $s_i$ and $s_j$ respectively.

We can write:
%\[\log \tilde{g}_{ij}q_{\neg \tilde{g}_{ij}} \propto \log \left(\frac{\theta_i}{1-\theta_i} \right) s_i + \log \left(\frac{\theta_j}{1-\theta_j} \right) s_j + \sum_{i=1}^{K}  \log \left(\frac{\lambda_i}{1-\lambda_i} \right) +  \sum_{i'=1}^{K}\sum_{\substack{j'=1 \\ i', j' \neq i, j}}^{i'-1}  \log \left(\frac{\theta_i}{1-\theta_i} \right) s_i + \log \left(\frac{\theta_j}{1-\theta_j} \right) s_j\]

\[\log \tilde{g}_{ij}(s_i, s_j)q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j) \propto \xi_{ji} s_i + \xi_{ij} s_j + \eta_{i, \neg s_j} s_i + \eta_{j, \neg s_i} s_j\]

Simplifying:

\[\log \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j) \propto \left(\xi_{ji}+ \eta_{i, \neg s_j} \right) s_i + \left(\xi_{ij} + \eta_{j, \neg s_i} \right) s_j \]

%\[
%\begin{array}{l}
% \log \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \log \left(\frac{\theta_i}{1-\theta_i} \right) s_i + \log \left(\frac{\theta_j}{1-\theta_j} \right) s_j\\
%\tab \tab \tab \tab \tab + \sum_{i'=1}^{K}  \log \left(\frac{\lambda_{i'}}{1-\lambda_{i'}} \right) s_{i'} \\
%\tab \tab \tab \tab \tab +  \sum_{i'=1}^{K}\sum_{\substack{j'=1 \\ i', j' \neq i, j}}^{i'-1}  \log \left(\frac{\theta_{i'}}{1-\theta_{i'}} \right) s_{i'} + \log \left(\frac{\theta_{j'}}{1-\theta_{j'}} \right) s_{j'}
%\end{array}
%\]

%Only including terms with $s_i$ and $s_j$:
%
%\[
%\begin{array}{l}
% \log \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \left((K-1) \log \left(\frac{\theta_i}{1-\theta_i} \right) + \log \left(\frac{\lambda_i}{1-\lambda_i} \right)\right) s_i\\
%\tab \tab \tab \tab \tab +\left(
%(K-1) \log \left(\frac{\theta_j}{1-\theta_j} \right) + \log \left(\frac{\lambda_j}{1-\lambda_j} \right)\right) s_j
%\end{array}
%\]

Thus, the first moments:

\[\mathbb{E}_{s_i}\left[\sum_{s_j \in \{0, 1\}}\tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j)\right] = \frac{1}{1+\exp{\left(-\left(\xi_{ji} + \eta_{i, \neg s_j} \right)}\right)}\]

and

\[\mathbb{E}_{s_j}\left[\sum_{s_i \in \{0, 1\}}\tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j)\right] = \frac{1}{1+\exp{\left(-\left(\xi_{ij} + \eta_{j, \neg s_i} \right)}\right)}\]

Moreover:

\[\log g_{ij}(s_i, s_j) qq_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j) \propto W_{ij} s_i s_j
 + \eta_{i, \neg s_j} s_i + \eta_{j, \neg s_i} s_j\]

%\[
%\begin{array}{l}
% \log g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \frac{- \mu_i^T \mu_j}{\sigma^2} s_i s_j\\
%\tab \tab \tab \tab \tab + \sum_{i'=1}^{K}  \log \left(\frac{\lambda_{i'}}{1-\lambda_{i'}} \right) s_{i'} \\
%\tab \tab \tab \tab \tab +  \sum_{i'=1}^{K}\sum_{\substack{j'=1 \\ i', j' \neq i, j}}^{i'-1}  \log \left(\frac{\theta_{i'}}{1-\theta_{i'}} \right) s_{i'} + \log \left(\frac{\theta_{j'}}{1-\theta_{j'}} \right) s_{j'}
%\end{array}
%\]

%Only including terms with $s_i$ and $s_j$:
%
%\[
%\begin{array}{l}
% \log g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \frac{- \mu_i^T \mu_j}{\sigma^2} s_i s_j\\
%\tab \tab \tab \tab \tab + \left((K-2)\log \left(\frac{\theta_i}{1-\theta_i} \right) + \log \left(\frac{\lambda_i}{1-\lambda_i} \right)\right) s_i \\
%\tab \tab \tab \tab \tab +  \left((K-2)\log \left(\frac{\theta_j}{1-\theta_j} \right) + \log \left(\frac{\lambda_j}{1-\lambda_j} \right)\right) s_j
%\end{array}
%\]
%
%We can write this as:
%
%\[
%\log g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto W_{ij} s_i s_j+ A_{s_i} s_i + B_{s_j} s_j
%\]
%
%where:
%
%\[A_{s_i} = (K-2)\log \left(\frac{\theta_i}{1-\theta_i} \right) + \log \left(\frac{\lambda_i}{1-\lambda_i} \right)\]
%
%and
%
%\[B_{s_j} = (K-2)\log \left(\frac{\theta_j}{1-\theta_j} \right) + \log \left(\frac{\lambda_j}{1-\lambda_j} \right)\]

To derive the first moment for $ g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j)$ with respect to $s_i$, we first marginalise out $s_j$:

\[ \sum_{s_j \in \{0, 1\}} g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \exp\left( W_{ij} s_i + \eta_{i, \neg s_j} s_i + \eta_{j, \neg s_i}\right) +  \exp\left(\eta_{i, \neg s_j} s_i \right)
\]

Thus, the first moment:

\[\mathbb{E}_{s_i}\left[\sum_{s_j \in \{0, 1\}}g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j)\right] = \frac{\exp\left( W_{ij} + \eta_{i, \neg s_j}  + \eta_{j, \neg s_i}\right) +  \exp\left(\eta_{i, \neg s_j}\right)}{\left[\exp\left( W_{ij} + \eta_{i, \neg s_j}  + \eta_{j, \neg s_i}\right) +  \exp\left(\eta_{i, \neg s_j}\right)\right]+\left[\exp\left( \eta_{j, \neg s_i}\right) +  1\right]}\]

Simplifying:

\[\mathbb{E}_{s_i}\left[\sum_{s_j \in \{0, 1\}}g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j)\right] = \frac{\exp\left(\eta_{i, \neg s_j}\right)\left(\exp\left( W_{ij}  + \eta_{j, \neg s_i}\right) + 1\right)  }{\left[\exp\left(\eta_{i, \neg s_j}\right)\left(\exp\left( W_{ij}  + \eta_{j, \neg s_i}\right) + 1\right)  \right]+\left[\exp\left( \eta_{j, \neg s_i}\right) +  1\right]}\]

Similarly:

\[\mathbb{E}_{s_j}\left[\sum_{s_i \in \{0, 1\}}g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j)\right] = \frac{\exp\left(\eta_{j, \neg s_i}\right)\left(\exp\left( W_{ij}  + \eta_{i, \neg s_j}\right) + 1\right)  }{\left[\exp\left(\eta_{j, \neg s_i}\right)\left(\exp\left( W_{ij}  + \eta_{i, \neg s_j}\right) + 1\right) \right]+\left[\exp\left( \eta_{i, \neg s_j}\right) +  1\right]}\]

By setting:

\[\mathbb{E}_{s_i}\left[\sum_{s_j \in \{0, 1\}}\tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j)\right] = \mathbb{E}_{s_i}\left[\sum_{s_j \in \{0, 1\}}g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j)\right]\]

and

\[\mathbb{E}_{s_j}\left[\sum_{s_i \in \{0, 1\}}\tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j)\right] = \mathbb{E}_{s_j}\left[\sum_{s_i \in \{0, 1\}}g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(s_i, s_j)\right]\]

we can solve for the parameters of $\tilde{g}_{ij}(s_i, s_j)$ with moment matching:

\[\frac{1}{1+\exp{\left(-\left(\xi_{ji} + \eta_{i, \neg s_j} \right)}\right)}} = \frac{\exp\left(\eta_{i, \neg s_j}\right)\left(\exp\left( W_{ij}  + \eta_{j, \neg s_i}\right) + 1\right)  }{\left[\exp\left(\eta_{i, \neg s_j}\right)\left(\exp\left( W_{ij}  + \eta_{j, \neg s_i}\right) + 1\right)  \right]+\left[\exp\left( \eta_{j, \neg s_i}\right) +  1\right]}\]

Simplifying:

\[\exp\left( \eta_{j, \neg s_i}\right) +  1=\exp{\left(-\left(\xi_{ji} + \eta_{i, \neg s_j} \right)}\right)}\exp\left(\eta_{i, \neg s_j}\right)\left(\exp\left( W_{ij}  + \eta_{j, \neg s_i}\right) + 1\right)\]

\[\frac{\exp\left( \eta_{j, \neg s_i}\right) +  1}{ \exp\left( W_{ij}  + \eta_{j, \neg s_i}\right) + 1} =\exp{\left(-\xi_{ji}\right)}\]

Our parameter update:

\[\xi_{ji} = \log \left( \frac{1+\exp\left( W_{ij}  + \eta_{j, \neg s_i}\right)}{1+\exp\left( \eta_{j, \neg s_i}\right)} \right)\]

Similarly:

\[\xi_{ij} = \log \left( \frac{1+\exp\left( W_{ij}  + \eta_{i, \neg s_j}\right)}{1+\exp\left( \eta_{i, \neg s_j}\right)} \right)\]

\subsection*{(c)}

Our message passing approximations:

\[\exp(\eta_{ij} s_i) = \tilde{f}_{i}(s_{i}) \prod_{k\in ne(i), k\neq j}^{K} \mathcal{M}_{k \rightarrow i} \]

\[\exp(\eta_{ji} s_j) = \tilde{f}_{j}(s_{j}) \prod_{k\in ne(j), k\neq i}^{K} \mathcal{M}_{k \rightarrow j}\]

where each message $\mathcal{M}_{j \rightarrow i}$ has a factored approximation:

\[\mathcal{M}_{k \rightarrow i} = \exp(\eta_{ki} s_k) \]

because each site $\tilde{g}_{jk}(s_j s_k)$ is approximated as a product of two messages $\mathcal{M}_{j \rightarrow k}\mathcal{M}_{k \rightarrow j}$, each a Bernoulli.

Thus, the natural parameters of the messages are updated with:

\[\eta_{ij} = b_i + \sum_{k\in ne(i), k\neq j}^{K} \eta_{ki} \]

The summation of the natural parameters of the singleton factor for node $i$ with the natural parameters of messages from all the neighbouring nodes.

This leads to a loopy BP algorithm because the nodes are fully connected (i.e. every node is the neighbour of all other nodes). Thus, we cannot simply move from one end of the graph to the other like BP for tree structured graphs.

\subsection*{(d)}

We can use automatic relevance determination (ARD) as a hyperparameter method to select relevant features

Place prior on $\sigma^2$ and optimise with respect to the distributions would cause some to diverge and only relevant latent dimensions will remain.
This gives us a value for $K$, the number of latent factors that haven't diverged.

\newpage
\section*{Question 6}

\begin{figure}[h]
\centering
\includegraphics[scale=0.4]{outputs/q6/all-latent-factors}
\caption{Latent factors learned with EP/Loopy-BP}
\label{fig:6-latent-factors}
\end{figure}

\newpage
The Python code for the Boltzmann machine:
\lstinputlisting[language=Python]{src/models/boltzmann_machine.py}

\newpage
The Python code for message passing:
\lstinputlisting[language=Python]{src/models/message_passing.py}

\newpage
The rest of the Python code for question 6:
\lstinputlisting[language=Python]{src/solutions/q6.py}

\newpage
\section*{Appendix 1: constants.py}
\lstinputlisting[language=Python]{src/constants.py}
\newpage
\section*{Appendix 2: main.py}
\lstinputlisting[language=Python]{main.py}
\end{document}